Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-31 analysis/paper/paper.Rmd
add text of metric data to discussion section

<!-- a few lines to compare metric to shape... main point to make clear is the substantive result about the main research questions -->

>>>>

When we examine metric data for six measurements and two ratio values using CV tests (\table here), it shows there is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter before and after the European presence. After the arrival of the Chinese, there is only one metric attribute, the rim diameter, showing significant difference compared with post-European period (p-value less than 0.05). Although these results shows the differences between different contact phases that corresponds to part of the results of shape variable, elliptic Fourier analysis captures more subtle differences in a substantially way reflected by overall shapes of pottery, not only limited to linear aspects.

### 2020-01-31 analysis/paper/scraps.Rmd
add text of metric data to discussion section

>>>>

#--------------original metric data analysis------------
Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%)  before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). <!-- these numeric values in this paragraph should come from inline code, not copy-pasted in like this -->
<!-- You need to put the code that computes the MSLRT results right in here, in the Rmd -->

```{r metricCVstest}
metric_cvs_with_phase <- readr::read_csv(here::here("analysis", "data", "raw_data", "Kwl_p_metrics_cv_test.csv"))
metric_cvs_with_phase_tidy <-
  metric_cvs_with_phase %>% 
  mutate_at(c("pre-e", "post-e", "ch-con", "MSLRT", "p_value"), round, 4) %>% 
  rename(`pre-European` = "pre-e", `after European` = "post-e", Chinese = "ch-con")
knitr::kable(metric_cvs_with_phase_tidy,
             caption = "Coefficient of Variation test for metric attributes by phases")
```

#------------revised metric data analysis---------------

Table \@ref(tab:metricCVstest-pre-post) and \@ref(tab:metricCVstest-post-chi) shows the results of Asymptotic test [@Feltz1996] for equality of CVs for for six measurements and two ratio values. There is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter before and after the European presence. The CV values show that the variation of rim diameter is small (6.8%) before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). After the arrival of the Chinese, there is only one metric attribute, the rim diameter, showing significant difference from post-European period (p-value less than 0.05).

```{r metricCVstest-pre-post, eval=FALSE}
# read metric data
kwl_p_ave <- readr::read_csv(here::here("analysis", "data", "raw_data",
                                 "KWL-pottery-metrics-tidy-ave.csv"))
library(cvequality)
library(tidyverse)
# select metric attributes and tidy data for computing CV
kwl_p_metrics_long <-
  kwl_p_ave %>%
  mutate(dia_rim_body = Dia_Rim_ave/Dia_Body_ave,
         thi_rim_body = Thick_Rim_ave/Thick_Body_ave) %>%
  select(period,
         Dia_Rim_ave,
         Dia_Neck_ave,
         Dia_Body_ave,
         Thick_Rim_ave,
         Thick_Neck_ave,
         Thick_Body_ave,
         dia_rim_body,
         thi_rim_body,
         # H_Rim
  ) %>%
  rename(`Rim diameter (mm)` = Dia_Rim_ave,
         `Neck diameter (mm)` = Dia_Neck_ave,
         `Body diameter (mm)` = Dia_Body_ave,
         `Rim thickness (mm)` = Thick_Rim_ave,
         `Neck thickness (mm)` = Thick_Neck_ave,
         `Body thickness (mm)` = Thick_Body_ave,
         `Ratio of Rim/Body diameter` = dia_rim_body,
         `Ratio of Rim/Body thickness` = thi_rim_body
  ) %>%
  gather(variable, value, -period) %>%
  filter(!is.na(period),
         !is.na(value))
# filter pre-e and post-e periods
kwl_p_metrics_long_pre_post <-
  kwl_p_metrics_long %>%
  filter(!period == "ch-con")
# filter post-e and chi periods
kwl_p_metrics_long_post_chi <-
  kwl_p_metrics_long %>%
  filter(!period == "pre-e")
#------------------------------------------CV test for pre-e and post-e pair
kiwulan_pottery_metrics_cvs_plot_cv_test_asym_pre_post <-
  kwl_p_metrics_long_pre_post %>%
  nest(-variable) %>%  # or nest(period, value) %>%
  mutate(asymptotic_test = map(data, # change from mslr_test to asymptotic_test
                               ~bind_cols(asymptotic_test (
                                 .x$value,
                                 .x$period)))) %>%
  unnest(asymptotic_test)
# compute CV
kwl_p_cvs_pre_post <-
  kwl_p_metrics_long_pre_post %>%
  group_by(period, variable) %>%
  summarise(cvs = raster::cv(value, na.rm = TRUE))
# join p-values to CVs
kwl_p_cvs_and_pvalues_pre_post <-
  kwl_p_cvs_pre_post %>%
  left_join(kiwulan_pottery_metrics_cvs_plot_cv_test_asym_pre_post) %>%
  mutate(significant = ifelse(p_value <= 0.05, 'yes', 'no')) %>%
  mutate(variable = factor(variable, ordered = TRUE)) %>%
  arrange(variable)
# making table for CV
kwl_p_cvs_table_pre_post <-
  kwl_p_cvs_and_pvalues_pre_post %>%
  select(-data, -significant) %>%
  pivot_wider(names_from = period, values_from = cvs) %>%
  mutate(variable = factor(variable, levels = c("Rim thickness (mm)",
                                                "Neck thickness (mm)",
                                                "Body thickness (mm)",
                                                "Ratio of Rim/Body thickness",
                                                "Rim diameter (mm)",
                                                "Neck diameter (mm)",
                                                "Body diameter (mm)",
                                                "Ratio of Rim/Body diameter"))) %>%
  arrange(variable) %>%
  select(variable, `pre-e`, `post-e`, `D_AD`, `p_value`) %>%
  mutate_at(c("pre-e", "post-e", "D_AD", "p_value"), round, 4)
knitr::kable(kwl_p_cvs_table_pre_post,
             caption = "Coefficient of Variation for metric attributes before and after European presence using Asymptotic test [@Feltz1996]")
```

```{r metricCVstest-post-chi, eval=FALSE}
#-----------------------------------CV test for post-e and ch-con pair
kiwulan_pottery_metrics_cvs_plot_cv_test_asym_post_chi <-
  kwl_p_metrics_long_post_chi %>%
  nest(-variable) %>%  # or nest(period, value) %>%
  mutate(asymptotic_test = map(data, # change from mslr_test to asymptotic_test
                               ~bind_cols(asymptotic_test (
                                 .x$value,
                                 .x$period)))) %>%
  unnest(asymptotic_test)
# compute CV
kwl_p_cvs_pre_post <-
  kwl_p_metrics_long_post_chi %>%
  group_by(period, variable) %>%
  summarise(cvs = raster::cv(value, na.rm = TRUE))
# join p-values to CVs
kwl_p_cvs_and_pvalues_post_chi <-
  kwl_p_cvs_pre_post %>%
  left_join(kiwulan_pottery_metrics_cvs_plot_cv_test_asym_post_chi) %>%
  mutate(significant = ifelse(p_value <= 0.05, 'yes', 'no')) %>%
  mutate(variable = factor(variable, ordered = TRUE)) %>%
  arrange(variable)
# making table for CV
kwl_p_cvs_table_post_chi <-
  kwl_p_cvs_and_pvalues_post_chi %>%
  select(-data, -significant) %>%
  pivot_wider(names_from = period, values_from = cvs) %>%
  mutate(variable = factor(variable, levels = c("Rim thickness (mm)",
                                                "Neck thickness (mm)",
                                                "Body thickness (mm)",
                                                "Ratio of Rim/Body thickness",
                                                "Rim diameter (mm)",
                                                "Neck diameter (mm)",
                                                "Body diameter (mm)",
                                                "Ratio of Rim/Body diameter"))) %>%
  arrange(variable) %>%
  select(variable, `post-e`, `ch-con`, `D_AD`, `p_value`) %>%
  mutate_at(c("post-e", "ch-con", "D_AD", "p_value"), round, 4)
knitr::kable(kwl_p_cvs_table_post_chi,
             caption = "Coefficient of Variation for metric attributes between European and Chinese presence using Asymptotic test [@Feltz1996]")
```
